---
title: "How to Add a Fitted Line to a Scatterplot in R"
author: "Andreu"
date: '2022-08-31'
output: html_document
---

```{r setup , echo = FALSE}
knitr::opts_chunk$set(echo = TRUE , warning = FALSE)
```

## HOW TO ADD A FITTED LINE TO A SCATTERPLOT IN R

This basic example code is based on the following **data**:

```{r data}
set.seed(123)                        # Set seed for reproducibility
x <- rnorm(30)                       # Create x variable
y <- x + rnorm(30)                   # Create correlated y variable
df <- data.frame(x , y)              # Create dataframe
```

The **package** that we will be using for this example is ggplot2:

```{r package}
#install.packages("ggplot2")
library(ggplot2)
```

First, we have to create the **scatter plot** that we will want to fit:

```{r scatterplot}
p <- ggplot(data = df , aes(x = x , y = y)) +
    geom_point()
```

Our scatter plot looks as follows:

```{r p , echo = FALSE}
p
```

Then, we add the **fitted line** with the function called ```geom_smooth()```:

```{r fitted 1 , message = FALSE}
p + geom_smooth()
```

After doing that, you can try to change some of the options of this function, such as ```method``` or ```formula```. Some of the methods available are:

    - "loess" -> this is the method used by default.
    - "lm" -> fits a linear model, giving the best line of it.
    - "rlm" -> works like "lm", but uses a robust fitting algorithm so that outliers don't affect the fit as much. It's part of the MASS package, so remember to load it first.

```{r fitted 2}
p + geom_smooth(method = "lm", formula = y ~ poly(x,3))
```

Finally, you can also experiment with the function ```stat_smooth``` and combining it with ```geom_smooth``` in infinite ways:

```{r fitted 3 , message = FALSE}
p + geom_smooth() + stat_smooth(se = TRUE , level = 0.7 , span = 0.6)
```

```{r fitted 4 , message = FALSE}
p + geom_smooth(aes(colour = "loess") , method = "loess", se = FALSE) + 
    geom_smooth(aes(colour = "lm")    , method = "lm"   , se = FALSE)
```






